\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-10-24 icra-2021.tex
implemented all latest Hector's suggestions

The proposed algorithm alters the energy consumption dynamically by means of mission-specific parameters: the Quality of Service (QoS) of the computations, and the trajectory-explicit equations (TEEs) adjustments. In the remainder of the paper, we strictly adopt the following notation.  We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and to the current trajectory as TEE. Our goal is a mission extension by optimizing both computations and adjustments as the UAV flies and its batteries drain. First, the algorithm optimizes computations requiring the UAV to include robot operating system (ROS) nodes. Then, it optimizes adjustments--a way to alter the trajectory--and guides the UAV using a vector field~\cite{de2017guidance} that converges smoothly to such trajectory. It relies on the assumptions of the mission being \emph{periodic} and \emph{uncertain}. The periodicity is directly observed, by e.g., the UAV flying in repetitive patterns, and the uncertainty accounts for the environmental interference with e.g., a fixed-wing UAV drifting due to windy weather. It addresses the periodicity modeling the energy with Fourier analysis--being the mission periodic, we expect the energy to evolve also periodically--and the uncertainty with a state estimator. It output the controls sequence (computations and adjustments) using robust output feedback model predictive control (MPC).

In the spirit of reducing costs, and resources, we showcase the algorithm using the problem of dynamic mission planning for a precision agriculture fixed-wing UAV. Such a scenario is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The mission consists of a UAV flying in circles and lines covering a polygon, detecting obstacles using a convolutional neural network (CNN), and informing grounded mobile robots employed for future harvesting--a survey mission optimized for the craft's dynamics. The algorithm plans the mission controlling the processing rate and the radius of the circles, affecting the distance between the lines. Data indicates a potential extension of up to 13 minutes over an hour by merely switching to the lowest computations. Figure~\ref{fig:overview} shows an initial slice of such a mission. The UAV first heads to the circular TEE $\varphi_0$, and later shifts to TEE $\varphi_0^+$, as e.g., windy weather required the adjustment of the control sequence to avoid battery depletion.

@@

Let a mission's stage be a generic continuous twice differentiable TEE $\varphi$ and a set of tasks $\psi_1,\dots,\psi_{\sigma}$. Moreover, let $t_f$ be the final time, $[\,\cdot\,]$ the set $\{0,1,\dots,\,\cdot\,\}$, $[\,\cdot\,]^+$ the set $\{0\}/[\,\cdot\,]$, and $\underline{\,\cdot\,},\overline{\,\cdot\,}$ the lower and upper bound of $\,\cdot\,$. If $\,\cdot\,$ is a value, the bounds are user defined and retrieved from a lookup table; if a set, they correspond to minimum and maximum of the set.

\begin{defn}[Stage and mission]\label{def:mission}
  At a time instant $k\in[0,t_f)$, the \emph{stage} is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \mathcal{M}_k:=\{(\varphi_k&(\,\cdot\,),\psi_1(\,\cdot\,),...,\psi_\sigma(\,\cdot\,))\mid\\\exists\,\,&\varphi_k(\,\cdot\,)\in\mathbb{C}_k,\,\\&\,\psi_i(\,\cdot\,)\in\mathbb{S}_{k,i}\,\forall i\in[\sigma]^+\},
  \end{split}\end{equation}
  where $\mathbb{C}_k:=[\underline{c}_k,\overline{c}_k]\subseteq\mathbb{R}$ is the TEEs set, and $\mathbb{S}_{k,i}:=[\underline{s}_{k,i},\overline{s}_{k,i}]\subseteq\mathbb{Z}_{\geq 0}$ the $i$-th task QoS set. The parameters of the TEE $\varphi_k$ are defined in Subsection~\ref{sec:model}, of the tasks $\psi_1,\dots,\psi_\sigma$ in~\ref{sec:computations-model}.

  The \emph{mission} is the union of all the stages, along with a function $\lambda:\mathbb{R}^2\rightarrow\mathcal{M}_k$ which maps a point $\mathbf{p}\in\mathbb{R}^2$ of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$ to a specific stage. If for simplicity the system is sampled discrete-time
  \begin{equation}
    \mathcal{M}:=(\lambda(\mathbf{p}),\bigcup_{i\in[t_f]}{\mathcal{M}_i}),
  \end{equation}
  the algorithm inputs $\mathcal{M}$, initial position, and energy coefficients guess, and outputs the position, the instantaneous energy consumption, and the controls sequence--an action performed evolving the mission state.
\end{defn}

@@

The set
\begin{equation}\label{eq:area}
  \mathcal{P}_k:=\{\mathbf{p}_k\mid\varphi_k(\mathbf{p}_k,c_{k,1},\dots,c_{k,\rho})\in\mathbb{C}_k\},
\end{equation}
delimits the area where the $k$-th TEE $\varphi_k:\mathbb{R}^2\times\mathbb{R}^{\rho}\rightarrow\mathbb{R}$ is free to evolve using $\rho$ adjustments $\mathbf{c}_k:=c_{k,1},...,c_{k,\rho}$ (gray area in Figure~\ref{fig:overview}), being the TEE satisfied for all the approaching points $\varphi_k\rightarrow\mathbb{C}_k$.

The algorithm uses the concept to select the adjustments s.t. $\varphi_k(\mathbf{p}_k,\mathbf{c}_k^0)$ has the highest energy value. It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the vector field of $\varPhi:=\varphi_k(\mathbf{p}_k,\mathbf{c}_k^0)$, deriving the direction to follow--the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is the gradient, $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind $w\in\mathbb{R}$.

The algorithm models the energy using as state energy coefficients $\mathbf{q}\in\mathbb{R}^j$ derived from Fourier analysis (the meaning of $j$ is clarified to the reader shortly) and decompose such evolution in the energy due to the trajectory, and computations--an approach adapted from our earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing UAV~\cite{seewald2020mechanical}.

@@

Let us consider a Fourier series $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of energy consumption modeling of the mission
\begin{equation}\label{eq:fourier}
  h(t)=\sum_{n=0}^{r}{a_n\cos\left({nt/\xi}\right)+b_n\sin{\left({nt/\xi}\right)}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ the Fourier series coefficients.

Suppose uncertainty in the form of $\mathbf{w}_k\in\mathbb{R}^j,v_k\in\mathbb{R}$ accounting for the unknown state and output is absent. The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear discrete time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k\\
  y_k&=C\mathbf{q}_k+v_k
\end{cases},\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ mimics the original Fourier series coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & \frac{n}{\xi} \\ -\frac{n^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, and $C\in\mathbb{R}^j$ is the output matrix. In matrix $A$, the first value is one, $A_n$ is later on the diagonal, and zero in the remainder.

The control $\mathbf{u}$ along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}  
  \mathbf{u}_k&=\begin{bmatrix}g(\mathbf{s}_k)-g(\mathbf{s}_{k-1}) & \mathbf{c}_k-\mathbf{c}_{k-1}\end{bmatrix}^T,\\
  B&=\left[\begin{array}{cccc}
    1& \omega_{k,1}& \cdots& \omega_{k,\rho}  \\
     &            0&       &  \\
     &             & \ddots&  \\
     &             &       & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{s}_k,g(\,\cdot\,)$ are defined in Subsection~\ref{sec:computations-model}, $\mathbf{u}_k\in\mathbb{R}^l$ is the control given $l:=1+\rho$, and $B\in\mathbb{R}^{j\times l}$. Moreover, the first item is one, while the others on the first row are gain factors $\omega_{k}\in\mathbb{R}$, quantifying the contribution of a given adjustment to the instantaneous energy consumption.

Note that the controls sequence $\mathbf{u}_k^a:=(\mathbf{c}_{k},\mathbf{s}_{k})$--an output of the algorithm--differs from the nominal control in Equation~(\ref{eq:state-perf}) $\mathbf{u}_k(\mathbf{u}_k^a,\mathbf{u}_{k-1}^a)$.

The energy evolution analysis necessitates the following realistic assumption.
\begin{assm}[Energy evelution periodicity]\label{assm:periodic} 
Given two time instants $k_1,k_2\in[t_f]$ s.t. $k_1>k_2$ and a constant value $n\in\mathbb{R}_{> 0}$, there exist an arbitrary constant displacement $e\in\mathbb{R}$
\begin{equation}
  \abs{y_{k}-y_{k+n}}= e\,\,\,\forall k\in[k_1,k_2].
\end{equation}
\end{assm}

@@

Equation~(\ref{eq:state-control}) accounts for the energy due to the computations. The energy due to the adjustments is merely a linear combination of the gain factor and the adjustment. Nevertheless, the change updates the path which will hence affect the reading from the sensors and adjust the energy coefficients. For instance, a decrement in the adjustment radius of a circle when the TEE is a circle, adds a negative contribution, thus simulates the lowering of instantaneous energy consumption.

%In the case of the system behaving ideally (i.e., with no uncertainty), we expect a state (energy coefficients) evolving accordingly to its output (instantaneous energy consumption). An observation summarized in the following Lemma, where we use norm to compare two energy coefficients vectors.

%\begin{lem}[State, output proportionality]\label{lem:state-vs-energy}
%  Suppose the system of Equation~(\ref{eq:state-perf}) evolves with no uncertainty ($\mathbf{w}=\mathbf{0},v=0$) and Assumption~\ref{assm:periodic} holds. Then
%  \begin{equation}
%    \|\mathbf{q}_{k_1}\|\geq\|\mathbf{q}_{k_2}\|\iff y_{k_1}\geq y_{k_2}.
%  \end{equation}
%\end{lem}
%\begin{proof}
%  Suppose $y_{k_1}\geq y_{k_2}$ is true. Given $k=k_1, k+n=k_2$, $y_{k}=\alpha_{k,0}+\alpha_{k,1}+\dots+\alpha_{k,r}$. Moreover, the time evolution of the model in Equation~(\ref{eq:state-perf}) can be expressed $y_{k+n}=(\alpha_{k,0}+B\mathbf{u}_k+\dots+B\mathbf{u}_{k+n-1})+p_1(n)\alpha_{k,1}+\dots+p_r(n)\alpha_{k,r}$, where $p_r(n):=\prod_{i=1}^{n/2}{r^3/\xi^3}$ depends on the order of a given coefficient and $n$. 
%  Hence, $\alpha_{k,0}+\alpha_{k,1}+\dots+\alpha_{k,r}\geq (\alpha_{k,0}+B\mathbf{u}_k+\dots+B\mathbf{u}_{k+n-1})+p_1\alpha_{k,1}+\dots+p_r\alpha_{k,r}$.
%
%  The magnitudes are $\|\mathbf{q}_{k}\|=(\alpha_{k,0}^2+\alpha_{k,1}^2+\beta_{k,1}^2+\dots+\alpha_{k,r}^2+\beta_{k,r}^2)^{1/2}$ and $\|\mathbf{q}_{k+n}\|=((\alpha_{k,0}+B\mathbf{u}_k+\dots+B\mathbf{u}_{k+n-1})^2+p_1^2(n)\alpha_{k,1}^2+p_1^2(n)\beta_{k,1}^2+\dots+p_r^2(n)\alpha_{k,r}^2+p_r^2(n)\beta_{k,r}^2)^{1/2}$.
%  By still inspecting the Equation~(\ref{eq:state-perf}), the coefficient $p_r^2(n)\beta_{k,r}^2$ can be expressed as a function of coefficients $\alpha$, $p_r^2(n)\beta_{k,r}^2=$
%  
%  Hence $\alpha_{k,0}^2+\alpha_{k,1}^2+\beta_{k,1}^2+\dots+\alpha_{k,r}^2+\beta_{k,r}^2\geq (\alpha_{k,0}+B\mathbf{u}_k+\dots+B\mathbf{u}_{k+n-1})^2+p_1^2\alpha_{k,1}^2+p_1^2\beta_{k,1}^2+\dots+p_r^2\alpha_{k,r}^2+p_r^2\beta_{k,r}^2$ and $()$.
%
%\end{proof}
% Lol wrong... Doesn't make sense this lemma at all...
%
\subsection{Energy evolution due to computations}
\label{sec:computations-model}

The energy cost of the computations (set of tasks being executed on an embedded board on the UAV) is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}, that measures software configurations empirically and builds a computational energy consumption model. Specifically, the tool builds a linear interpolation, one per each task. It requires the user to implement the mission as a ROS system with one or more ROS nodes changing the computational load by node-specific ROS parameters. A way to simulate the change of the computations.

The plan $\mathcal{M}$ contains a set of ordered lists with $\sigma$ tasks each (recall Definition~\ref{def:mission}). These tasks are simulated by $\sigma$ ROS nodes $\Psi(\mathbf{s}_k):=\left(\psi_1(s_{k,1}),\dots,\psi_\sigma(s_{k,\sigma})\right)$, in fact they input the desired and output the actual computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_k:=\left\{s_{k,1},\dots ,s_{k,\sigma}\mid \psi_i(s_{k,i})\in\mathbb{S}_{k}\,\forall i\in[\sigma]^+\right\},
\end{equation}
where $s_{k,i}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $i$-th computation, $\mathbf{s}_k\in\mathbb{S}_k\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the set of $\sigma$ computations at time $k$ given $\mathbb{S}_k:=\bigcup_{i\in[\sigma]^+}{\mathbb{S}_{k,i}}$.

Let us further define $g:\mathbb{Z}_{\geq 0}\times\mathbb{Z}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous computational energy consumption value obtained interrogating \powprof{}
\begin{equation}\label{eq:energy-comp}\begin{split}
  y_k^s:=g\left(\text{QoS}_{k,1},\dots,\text{QoS}_{k,\sigma}\right)=g\left(\mathbf{s}_{k}\right).\\
\end{split}\end{equation}
Moreover, $g(\{\emptyset\})=0$.

@@

Given an initial stage $\mathcal{M}_0$ (TEE $\varphi_0$, tasks $\Psi$, computations $\mathbf{s}_0\in\mathbb{S}_0$, and adaptations $\mathbf{c}_0\in\mathbb{C}_0$), the main purpose of the algorithm is to output a set of control sequences $\mathbf{u}^a:=\{\mathbf{u}_0^a,\mathbf{u}_1^a,\dots\}$ in a valid mission.

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{k-1}$ in $k\in[t_f]^+$ there exist a control sequence $\mathbf{u}_k^{a}$ that produce the next stage $\mathcal{M}_k$
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{(\mathbf{c}_{k},\mathbf{s}_{k})\mid& \,\exists\,\mathbf{c}_{k-1}\in\mathbb{C}_{k-1}, \mathbf{s}_{k-1}\in\mathbb{S}_{k-1}\text{ and}\\
    &(\varphi(\mathbf{p}_{k-1},\mathbf{c}_{k-1}),\Psi(\mathbf{s_{k-1}}))\in\mathcal{M}_{k-1}\\
    &\Longrightarrow(\varphi(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k}))\in\mathcal{M}_{k}\}.
  \end{split}\end{equation}
\end{defn}

Let us proof that if the mission is valid, the instantaneous energy consumption can be modeled as linear combination of the state from the Equation~(\ref{eq:state-perf}) for all $k\in[t_f]$.

>>>>

The proposed algorithm alters the energy consumption dynamically by means of mission-specific parameters: the Quality of Service (QoS) of the computations, and the trajectory-explicit equations (TEEs) adjustments. In the remainder of the paper, we strictly adopt the following notation.  We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and to the current trajectory as TEE. Our goal is a mission extension by optimizing both computations and adjustments as the UAV flies and its batteries drain. First, the algorithm optimizes computations requiring the UAV to include robot operating system (ROS) nodes. Then, it optimizes adjustments--a way to alter the trajectory--and guides the UAV using a vector field~\cite{de2017guidance} that converges smoothly to such trajectory. It relies on the assumptions of the mission being \emph{periodic} and \emph{uncertain}. The periodicity is directly observed, by e.g., the UAV flying in repetitive patterns, and the uncertainty accounts for the environmental interference with e.g., a fixed-wing UAV drifting due to windy weather. It addresses the periodicity modeling the energy with Fourier analysis--being the mission periodic, we expect the energy to approximately evolve also periodically--and the uncertainty with a state estimator. It output the control input (computations and adjustments) using robust output feedback model predictive control (MPC).

In the spirit of reducing costs, and resources, we showcase the algorithm using the problem of dynamic mission planning for a precision agriculture fixed-wing UAV. Such a scenario is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The mission consists of a UAV flying in circles and lines covering a polygon, detecting obstacles using a convolutional neural network (CNN), and informing grounded mobile robots employed for future harvesting--a survey mission optimized for the craft's dynamics. The algorithm plans the mission controlling the processing rate and the radius of the circles, affecting the distance between the lines. Figure~\ref{fig:overview} shows an initial slice of such a mission. The UAV first heads to a circular TEE with a given radius, which is later reduced, as e.g., windy weather required the adjustment of the control input to avoid battery depletion. Data indicates a potential extension of up to 13 minutes over an hour by merely switching to the lowest computations.

@@

The mission consists of different stages. At each stage the robot must follow a path and do some tasks.

Let the path at stage $i$ be characterized by a generic continuous twice differentiable TEE $\varphi_i:\mathbb{R}^2\times\mathbb{R}^\rho\rightarrow\mathbb{R}$ and the tasks by $\psi_1,\dots,\psi_{\sigma}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ functions. Moreover, let $[a]$ be the set $\{0,1,\dots,a\}$, $[a]^+$ the set $\{0\}/[a]$, and $\underline{c},\overline{c}$ the lower and upper bound of the parameter $c$ retrieved from a lookup table.

\begin{defn}[Stage and mission]\label{def:mission}
  A \emph{stage} $i$ at time instant $k$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \mathcal{M}_i:=\{(\varphi_i&(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho}),\\
      \psi_1&(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma}))\mid\\
      \exists\,\,\varphi&_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      \psi&_j(s_{i,j})\in\mathbb{S}_{i,j}\,\forall j\in[\sigma]^+\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the TEEs set, and $\mathbb{S}_{i,j}:=[\underline{s}_{i,j},\overline{s}_{i,j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th task QoS set. The parameters of the TEE $\varphi_i$ are defined in Subsection~\ref{sec:model}, of the tasks $\psi_1,\dots,\psi_\sigma$ in~\ref{sec:computations-model}.

  The \emph{mission} is the union of all the stages, along with a function $\lambda:\mathbb{R}^2\rightarrow\mathcal{M}_i$ which maps a point $\mathbf{p}\in\mathbb{R}^2$ of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$ to a specific stage. For simplicity the system is sampled discrete-time and $l$ is the total number of stages
  \begin{equation}
    \mathcal{M}:=(\lambda(\mathbf{p}),\bigcup_{i\in[l]}{\mathcal{M}_i}),
  \end{equation}
  the algorithm inputs $\mathcal{M}$, initial position, and energy coefficients guess, and outputs the position, the instantaneous energy consumption, and the control input--an action performed evolving the mission state.
\end{defn}

@@

The set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using $\rho$ adjustments $\mathbf{c}_i:=c_{i,1},...,c_{i,\rho}$ (gray area in Figure~\ref{fig:overview}), being the TEE satisfied for all the approaching points $\varphi_i\rightarrow\mathbb{C}_i$.

The algorithm uses the concept to select the adjustments s.t. $\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$ has the highest energy value. It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the vector field of $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, deriving the direction to follow--the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is the gradient, $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind $w\in\mathbb{R}$.

The algorithm models the energy using as state energy coefficients $\mathbf{q}\in\mathbb{R}^m$ derived from Fourier analysis (the meaning of $m$ is clarified to the reader shortly) and decompose such evolution in the energy due to the trajectory, and computations--an approach adapted from our earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing UAV~\cite{seewald2020mechanical}.

@@

Let us consider a Fourier series $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of energy consumption modeling of the mission
\begin{equation}\label{eq:fourier}
  h(k)=\sum_{i=0}^{r}{a_i\cos\left({ik/\xi}\right)+b_i\sin{\left({ik/\xi}\right)}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a,b\in\mathbb{R}$ the Fourier series coefficients.

Suppose uncertainty in the form of $\mathbf{w}_k\in\mathbb{R}^m,v_k\in\mathbb{R}$ accounting for the unknown state and output is absent. The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear discrete time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k\\
  y_k&=C\mathbf{q}_k+v_k
\end{cases},\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ mimics the original Fourier series coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_i=\begin{bmatrix}0 & \frac{i}{\xi} \\ -\frac{i^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^m$ given $m:=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the first value is one, $A_i$ is later on the diagonal, and zero in the remainder.

Suppose at time instant $k$ the mission reached stage $i$ and the control input is $\mathbf{u}_k^a=(\mathbf{c}_k,\mathbf{s}_k)$. The control along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}  
  \mathbf{u}_k&=\begin{bmatrix}g(\mathbf{s}_k)-g(\mathbf{s}_{k-1}) & \mathbf{c}_k-\mathbf{c}_{k-1}\end{bmatrix}^T,\\
  B&=\left[\begin{array}{cccc}
    1& \omega_{i,1}& \cdots& \omega_{i,\rho}  \\
     &            0&       &  \\
     &             & \ddots&  \\
     &             &       & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{s}_k,g(\mathbf{s}_k)$ are defined in Subsection~\ref{sec:computations-model}, $\mathbf{u}_k\in\mathbb{R}^n$ is the control given $n:=1+\rho$, and $B\in\mathbb{R}^{m\times n}$. Moreover, the first item is one, while the others on the first row are gain factors $\omega_{i}\in\mathbb{R}$, quantifying the contribution of a given adjustment to the instantaneous energy consumption.

Note that the control input $\mathbf{u}_k^a:=(\mathbf{c}_{k},\mathbf{s}_{k})$--an output of the algorithm--differs from the nominal control in Equation~(\ref{eq:state-perf}) $\mathbf{u}_k(\mathbf{u}_k^a,\mathbf{u}_{k-1}^a)$.

The energy evolution analysis necessitates the following realistic assumption.
\begin{assm}[Energy evelution periodicity]\label{assm:periodic} 
Given two time instants $k_1,k_2$ s.t. $k_1>k_2$ and a constant value $n\in\mathbb{Z}_{> 0}$, there exist an arbitrary constant displacement $e\in\mathbb{R}_{\geq 0}$
\begin{equation}
  \abs{y_{k}-y_{k+n}}= e\,\,\,\forall k\in[k_1,k_2].
\end{equation}
\end{assm}

@@

Equation~(\ref{eq:state-control}) accounts for the energy due to the computations. The energy due to the adjustments is merely a linear combination of the gain factor $\omega$ and the adjustment. Nevertheless, the change updates the path which will hence affect the reading from the sensors and adjust the state. For instance, a decrement in the adjustment radius of a circle when the TEE is a circle, adds a negative contribution, thus simulates the lowering of instantaneous energy consumption.

\subsection{Energy evolution due to computations}
\label{sec:computations-model}

The energy cost of the computations (set of tasks being executed on an embedded board on the UAV) is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}, that measures software configurations empirically and builds a computational energy consumption model. Specifically, the tool builds a linear interpolation, one per each task. It requires the user to implement the mission as a ROS system with one or more ROS nodes changing the computational load by node-specific ROS parameters.

The mission $\mathcal{M}$ contains a set of ordered lists with tasks (recall Definition~\ref{def:mission}). These tasks are simulated by ROS nodes $\Psi(\mathbf{s}_i):=\left(\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\right)$, in fact they input the desired and output the actual computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_i:=\left\{s_{i,1},\dots ,s_{i,\sigma}\mid \psi_j(s_{i,j})\in\mathbb{S}_{i}\,\forall j\in[\sigma]^+\right\},
\end{equation}
where $s_{i,j}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $j$-th desired computation at stage $i$, $\mathbf{s}_i\in\mathbb{S}_i\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the union of all QoS sets given $\mathbb{S}_i:=\bigcup_{j\in[\sigma]^+}{\mathbb{S}_{i,j}}$.

Let us further define $g:\mathbb{Z}_{\geq 0}\times\mathbb{Z}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ as the actual instantaneous computational energy consumption value obtained interrogating \powprof{}
\begin{equation}\label{eq:energy-comp}\begin{split}
  y_k^s:=g\left(\Psi(\mathbf{s}_i)\right)=g\left(\mathbf{s}_{k}\right).\\
\end{split}\end{equation}
Moreover, $g(\{\emptyset\})=0$.

@@

Given an initial stage $\mathcal{M}_0$ (TEE $\varphi_0$, tasks $\Psi$, computations $\mathbf{s}_0\in\mathbb{S}_0$, and adaptations $\mathbf{c}_0\in\mathbb{C}_0$), the main purpose of the algorithm is to output a set of control input sequence $\mathbf{u}^a:=\{\mathbf{u}_0^a,\mathbf{u}_1^a,\dots\}$ in a valid mission.

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{i-1},\,i\in[l]^+$ there exist a control input $\mathbf{u}_k^{a}$ that produce the next stage $\mathcal{M}_i$
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{(\mathbf{c}_{k},\mathbf{s}_{k})\mid&(\varphi_i(\mathbf{p}_{k-1},\mathbf{c}_{i-1}),\Psi(\mathbf{s_{i-1}}))\in\mathcal{M}_{i-1}\\
    &\Longrightarrow(\varphi_i(\mathbf{p}_{k},\mathbf{c}_{i}),\Psi(\mathbf{s}_{i}))\in\mathcal{M}_{i}\}.
  \end{split}\end{equation}
\end{defn}

Let us proof that if the mission is valid, the instantaneous energy consumption can be modeled as linear combination of the state from the Equation~(\ref{eq:state-perf}).

\begin{thm}[Periodic energy model]\label{thm:state-vs-energy}
  Consider the mission from Definition~\ref{def:mission}, the valid mission from~\ref{def:valid}. Assume Assumption~\ref{assm:periodic} holds, the model of Equation~(\ref{eq:state-perf}) behaves ideally ($\mathbf{w}=\mathbf{0},v=0$), the initial energy coefficients state $\mathbf{q}_0$ is $y_0^a/m$ for the first coefficient where $y_0^a\in\mathbb{R}_{>0}$ is an initial measurement\footnote{$y_0^a$ can be the initial measurement or the measurement from a previous instance of the algorithm}, $(1/2)y_0^a/m$ for all the others, and the mission is valid.
  Then, the instantaneous energy consumption $y_k$ is a linear combination of the state $\mathbf{q}_k$.
\end{thm}
\begin{proof}
  The proof is based on mathematical induction. 
  Base case: we proof that $y_0=y_0^a$. Recall the definition of the state in Equation~(\ref{eq:state-details}). The output is $y_{0}=\alpha_{0,0}+\alpha_{0,1}+\dots+\alpha_{0,r}=y_0^a/m+(1/2)y_0^a/m+\cdots+(1/2)y_0^a/m=y_0^a$.

  Induction step: by inspection of Equation~(\ref{eq:state-perf}), the output at instant $k$ can be expressed $y_{k}=(\alpha_{0,0}+B\mathbf{u}_0+\dots+B\mathbf{u}_{k-1})+p_1(k)\alpha_{0,1}+\dots+p_r(k)\alpha_{0,r}$, where $\forall t\in\mathbb{Z}_{\geq 2}$
  \begin{equation}\label{eq:proof-pr}
    p_r(t):=\begin{cases}
      \prod_{i=1}^{t/2}{r^3/\xi^3}&\text{ for even }t\\
      (r/\xi)\prod_{i=1}^{(t-1)/2}{r^3/\xi^3}&\text{ for odd }t
    \end{cases}.
  \end{equation}

  Suppose $k$ is even and the theorem holds up to $k$. Initial energy coefficients state $\mathbf{q}_0$ leads to $y_{k}=(y_0^a/m+B\mathbf{u}_0+\dots+B\mathbf{u}_{k-1})+p_1(k)(1/2)y_0^a/m+\dots+p_r(k)(1/2)y_0^a/m=y_{k}^a$.
